---
title: " "
author: " "
date: " "
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note:  

- Change the information in the yaml header above:  title, author, data.
- Make sure output is html_document.
- Once you are finished coding, run each chunk individually to make sure there are no errors.  (If necessary fix your code.) Once your code is error-free, click "knit" on the menu above. Your document should compile to HTML, provided that you have output set to "html_document."
- In the code chunk above ("setup") echo is set to TRUE.  This means that the code in your chunks will be displayed, along with the results, in your compiled document.
- Please delete these and other directions before compiling and submitting your assignment.

## Load packages and create empty data frame

```{r}
# Load packages
library(tidyverse) 

# Define n
n <- 1000

# Empty data frame for 1000 simulated days
d <- data.frame(day = seq(1:n),
                demand = NA,
                quantity = NA,
                price = NA,
                earnings = NA)

head(d)

```

Your goal in this project is to:

1. Fill in the dataset with random numbers drawn from the probability distributions defined in the case description---distributions for demand, quantity and price.
2. Use those numbers to compute earnings.
3. Summarize the earnings distribution to inform a business recommendation.

## Questions

Lightly comment your code and use pipes for readability.

Comment briefly on each of the questions, as directed.  Only the the final question requires a more lengthy (~ 2 paragraph) response.

### Q1


```{r}

```

### Q2


```{r}

```

### Q3

```{r}

```

### Q4

```{r}

```


### Q5

```{r}

```

### Q6


```{r}

```

## Q7

